Automated topic naming to support cross-project analysis of software maintenance activities

### 2011-03-24 msr/msr2011.tex
- [X] 4RQ2: What purpose does knowing how projects differ in       topics serve? Might be of use for a high-level view of the       system for managers, but authors need to justify           - response: we have addressed this in the text
Previous work, present term and terms     - [X] Previous work is written as if it should come at the end of       the paper. Too much terminology is used that has not yet been       introduced.           - response: we have addressed this issue and added             definitions to the undefined and clarified what our work             was.     - [X] should use

The idea of extracting higher-level \emph{concerns}, also known as \emph{concepts}, \emph{aspects} or \emph{requirements}, has been approached from documentation-based and repository-based perspectives.

Cleland-Huang and her colleagues have research on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
Their approach is similar to ours, as they mined keywords from NFR catalogues.
%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
We make cross-project comparisons, while they focused on a single project.
%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project. 
Another difference is that our data was far less structured than the requirements documents used in their study.
%There are several reasons we did not follow this route. 
%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
%Secondly, we wanted to compare across projects. 
%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
%A common taxonomy allows us to make cross-project comparison. 
%Although subject to the assumption that all projects conceive of these terms in the same way.
%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.

Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
They only used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
Mockus et al. validated their labels with system developers.
Since we study multiple projects, instead of a single project, these
kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).

@@

Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
project, similar to this, that identifies changes in quality requirements
in GNOME software projects; this approach was more exploratory, had
less validation, uses different word-lists, solely uses text-matching,
and does not leverage machine learning strategies. 
This research extends Hindle et al.~\cite{Hindle09ICSM} with labelled
topics, NFRs and new visualizations.

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
%\todo[inline]{more on intro ... why should the reader care about
 % this. It needs to show how the stuff in the intro (applications) is
 % possible}
%%% XXX NEIL I don't get this.. VVVVV
%\todo[inline]{fix section headings}
As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
In this section we demonstrate the value of labelled topic extraction
in addressing this issue. 
% more why
Labelled topics address \emph{why}  because they show why changes
occurred rather than \emph{how}. 
The \emph{how} of a change is the change itself, the purpose or
\emph{why} is what we are after.
We investigate the history of the two large-scale database systems that we studied. 
We use our technique to show the topic of development efforts over time in each project.
%We evaluated two research questions using the ground truth data that we annotated by hand:
We motivated our investigation with two research questions:
\begin{enumerate}
\item \emph{Do NFR frequencies change over time?} If a particular NFR was of more interest at one point in the life-cycle than another, this suggests that development activity shifted focus. For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, he or she might re-prioritize up-coming work items.
\item \emph{Do projects differ in their relative interest in NFRs?} A
  project manager, especially a systems-manager, would be interested in knowing whether a particular NFR, such as \emph{reliability}, was more important for one project than another. This could be to confirm the initial design goals, or to track the progress on that quarter's objectives.  
\end{enumerate}

>>>>

The idea of extracting higher-level \emph{concerns}, also known as
\emph{topics}, \emph{concepts}, \emph{aspects} or \emph{requirements},
has been approached from documentation-based and repository-based
perspectives.

Cleland-Huang and her colleagues have investigated on mining requirements documents for non-functional requirements (NFR) (quality requirements)~\cite{Cleland-Huang2006}. 
Their approach is similar to ours, as they mined keywords from NFR catalogues.
%One approach they tried was similar to this one, with keywords mined from NFR catalogues found in their previous work~\cite{chung99}. 
They demonstrated a recall of $80\%$ with precision of $57\%$ for the \emph{security} NFR, but could not find a reliable source of keywords for other NFRs. 
Instead, they developed a supervised classifier by using human experts to identify an NFR training set. 
Our research differs because we use a more comprehensive set of terms based on the taxonomy we chose.
In this paper we make cross-project comparisons instead of focusing
on a single project.
They relied on requirements documents instead of version control
histories that we use.
%The objective of Cleland-Huang's study was to identify new NFRs for system development, yet our objective was to recover those latent NFRs from commit-log messages of the project.

%There are several reasons we did not follow this route. 
%One, we believe we have a more comprehensive set of terms due to the taxonomy we chose. 
%Secondly, we wanted to compare across projects. 
%Their technique was not compared across different projects and the applicability of the training set to different corpora is unclear. 
%A common taxonomy allows us to make cross-project comparison. 
%Although subject to the assumption that all projects conceive of these terms in the same way.
%Thirdly, while the objective of Cleland-Huang's study was to identify new NFRs (for system development) our study assumes these NFRs are latent in the commit-log messages of the project.

Similarly, Mockus and Votta~\cite{Mockus00} studied a large-scale industrial change-tracking system. 
Mockus and Votta leveraged WordNet~\cite{Fellbaum1998}, an English-language ``lexical database'' that contains semantic relations between words, including common related forms (similar to word stemming), meronymy and synonymy.
They used WordNet for word roots as they felt the synonyms would be non-specific and cause errors, problems we also encountered. 
Mockus et al. validated their labels with system developers.
Since we study multiple projects, instead of a single project, these
kind of interviews are somewhat infeasible (particularly in the distributed world of open-source software).

@@

Finally, in Ernst et al.~\cite{ernst10refsq}, we describe our earlier
project, similar to this, that identifies changes in quality requirements
in GNOME software projects; this approach was more exploratory, had
less validation, uses different word-lists, solely uses text-matching,
and does not leverage machine learning strategies. 
Hindle et al.~\cite{Hindle09ICSM} propose a windowed method of topic
analysis, we extend this with labelled
topics, NFRs and new visualizations.

@@

\section{Understanding software maintenance activities} 
\label{sec:analysis}
% \todo[inline]{talk about this topic in one section}
% \todo[inline]{reference project name Nomen more}
%In order to test  our ROC values produced acceptable results, we now turn to
%\emph{applying} those techniques to understand software maintenance
%activities. 
%\todo[inline]{more on intro ... why should the reader care about
 % this. It needs to show how the stuff in the intro (applications) is
 % possible}
%%% XXX NEIL I don't get this.. VVVVV
%\todo[inline]{fix section headings}
As we mentioned in the introduction, a key issue in software maintenance is understanding \emph{why} a system has evolved the way it has~\cite{aranda09icse}. 
In this section we demonstrate the value of labelled topic extraction
in addressing this issue. 
% more why
Labelled topics address \emph{why}  because they show why changes
occurred rather than \emph{how}. 
The \emph{how} of a change is the change itself, the purpose or
\emph{why} is what we are after.
We investigate the history of the two large-scale database systems that we studied. 
We use our technique to show the topic of development efforts over time in each project.
%We evaluated two research questions using the ground truth data that we annotated by hand:
We motivated our investigation with two research questions:
\begin{enumerate}
\item \emph{Do NFR frequencies change over time?} If a particular NFR was of more interest at one point in the life-cycle than another, this suggests that development activity shifted focus. For example, if a developer expected to see a recent focus on \emph{reliability}, but instead \emph{usability} dominated, he or she might re-prioritize up-coming work items.
\item \emph{Do projects differ in their relative interest in NFRs?} A
  project manager, especially a systems-manager, would be interested
  in knowing whether a particular NFR, such as \emph{reliability}, was
  more important for one project than another. This could be to
  confirm the initial design goals, or to track the progress on that
  quarter's objectives.  The difference in NFR proportion is
  interesting because it implies a difference in focus between two projects.
\end{enumerate}
